Reduced MHD Notes

### 2025-01-27 Notes/Notes.tex
Edits
So many changes

\title{Reduced MHD Notes}
\author{Finny Valorz}
\date{June 2024}
\begin{document}
\maketitle

@@

More explicitly, we observe that $D_yf_0(x,y)$ is not invertible, failing to satisfy the constraint \eqref{constraint}. 
\begin{equation} \begin{split}
    D_yf_0(x,y) [\delta y] = \pth{\deriv{f_0^j}{y^i} \br{\delta y^i}} 
    &= \pmat{\deriv{}{r}f_0^r & \deriv{}{\bm{\nu}_\perp}f_0^r & \deriv{}{\beta_\para}f_0^r\\ \deriv{}{r}f_0^{\bm{\nu}_\perp} & \ddots & \vdots \\ \deriv{}{r}f_0^{\beta_\para} & \cdots & \deriv{}{\beta_\para}f_0^{\beta_\para}} 
    \bmat{\delta r \\ \delta \bm{\nu}_\perp \\ \delta \beta_\para} \\
    &= \pmat{0 & -\np \cdot & 0 \\ 0 & 0 & -\np \\ 0 & -\np \cdot & 0} 
    \bmat{\delta r \\ \delta \bm{\nu}_\perp \\ \delta \beta_\para}
    = \pmat{-\np \cdot \delta \bm{\nu}_\perp \\ -\np \delta \beta_\para \\ -\np \cdot \delta \bm{\nu}_\perp }
\end{split} \end{equation}
Remember that partial derivatives are treated as in \eqref{partial} as $\p f_0^j/\p y^i \br{\delta y^i} = D_{y^i} f_0^j \br{\delta y^i}$. Regardless of the non-zero entries, this block-matrix has a column of zeroes, so its determinant is $0$ and it is non-invertible. {\color{red}It is probably best to avoid the notion of determinant of an operator between infinite-dimensional spaces. Invertibility is a dimension-independent notion, but determinant is not. --JB} Intuitively, a fast-slow system ought to change on the fast timescale with a change in one of its fast variables. The system above is degenerate in the fast variable $r$, which suggests there are hidden slow variables among these fast variables.

@@

!!!Here is how I should have done it!!!
\begin{align*}
    \p_\tau\nu_\para &= -\ep\nu_\para\p_Z\nu_\para - \bm{\nu}_\perp\cdot \np\nu_\para \\ 
        &- \frac{1}{M_0^2\beta_0} \fr \pth{\beta_0\ep \p_ZP\pth{\frinv} + \ep^2\pth{1+\ep\beta_\para}\p_Z\beta_\para + \frac{1}{2}\ep^3\p_Z|\bm{\beta}_\perp|^2} \\ 
        &+ \frac{1}{M_0^2\beta_0} \fr\ep^2 \pth{(1+\ep\beta_\para)\p_Z\beta_\para + \bm{\beta}_\perp\cdot\np\beta_\para}. 
\end{align*}

@@

Equation \eqref{ndfaradayperp} gives an equation for the evolution of $\bm{\beta}_\perp = \np\Phi + \bm{e}_z\times\np\Psi$. 
\begin{align*} 
    \p_\tau\bm{\beta}_\perp = \p_Z \pth{\pth{1+\ep\beta_\para} \bm{\nu}_\perp - \ep\nu_\para \bm{\beta}_\perp} - \bm{e}_z \times \np \pth{\bm{e}_z \cdot \bm{\nu}_\perp \times \bm{\beta}_\perp}. 
\end{align*}
We use similar techniques to find $\Phi$ and $\Psi$ as we did for $\phi$ and $\psi$. The divergence of \eqref{ndfaradayperp} gives an expression for $\Phi$. 
\begin{align} \label{Phievolution} 
    \np^2\p_\tau\Phi = \np\cdot\p_\tau\bm{\beta}_\perp &= \p_Z \pth{\ep\np\beta_\para\cdot\bm{\nu}_\perp + (1+\ep\beta_\para) \np\cdot\bm{\nu}_\perp - \ep\np\nu_\para\cdot\bm{\beta_\perp} - \ep\nu_\para\np\cdot\bm{\beta_\perp}} \nonumber\\ 
    &= \p_Z\pth{\ep\Braket{\beta_\para, \phi} - \ep\br{\beta_\para, \psi} + (1+\ep\beta_\para) \np^2\phi - \ep\Braket{\nu_\para, \Phi} + \ep\br{\nu_\para, \Psi} -\ep\nu_\para\np^2\Phi} \nonumber\\ 
        &= \ep\Braket{\p_Z\beta_\para, \phi} + \ep\Braket{\beta_\para, \p_Z\phi} - \ep\br{\p_Z\beta_\para, \psi} - \ep\br{\beta_\para, \p_Z\psi} \nonumber\\ 
        &+ \ep\p_Z\beta_\para\np^2\phi + \pth{1+\ep\beta_\para}\np^2\p_Z\phi - \ep\Braket{\p_Z\nu_\para, \Phi} - \ep\Braket{\nu_\para, \p_Z\Phi} \nonumber\\ 
        &+ \ep\br{\p_Z\nu_\para, \Psi} + \ep\br{\nu_\para, \p_Z\Psi} - \ep\p_Z\nu_\para\np^2\Phi - \ep\nu_\para\np^2\p_Z\Phi. \text{\color{green}CHECKED} 
\end{align}
Note the second term in the equation of motion disappears because it involves the divergence of a curl.

@@

\subsection{$\nu_\para$ Evolution}
\begin{align} 
    \p_T\nu_\para &= \ep S_0 + \ep S_1 + \ep S_2 \text{, with}
\end{align}
\begin{align*}
    \ep S_0 &= -\ep^2\nu_\para \p_Z\nu_\para - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi}, \\ 
    \ep  S_1 &= -\beta_0\fr\p_Z\pi\pth{\frinv} - \cancel{\ep\p_Z\beta_\para} - \cancel{\ep^2q\p_Z\beta_\para} + \cancel{\ep^2\beta_\para\p_Z\beta_\para} \\ 
        &- \cancel{\ep^2\beta_\para\p_Z\beta_\para} - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 + O\pth{\ep^3}, \\ 
    \ep S_2 &= \cancel{\ep\p_Z\beta_\para} + \cancel{\ep^2 q\p_Z\beta_\para} - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} \\ 
        &- \ep\br{\beta_\para, \Psi} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi} + O\pth{\ep^3}. 
\end{align*}
Only one term changes in different regimes, namely 
\begin{align*}
    -\beta_0\fr\p_Z\pi\pth{\frinv} &= - \beta_0\ep\pi'(1) \p_Z\beta_\para + \beta_0\ep\pi'(1) \p_Zq \\ 
        &- \beta_0\ep^2\pi'(1) \pth{\cancel{q}-\beta_\para} \p_Z\beta_\para + \cancel{\beta_0\ep^2\pi'(1) \pth{q-\beta_\para} \p_Zq} \\ 
        &- \beta_0\ep^2\pi''(1) \beta_\para\p_Z \beta_\para + \beta_0\ep^2\pi''(1) \beta_\para\p_Zq + \beta_0\ep^2\pi''(1) q\p_Z\beta_\para - \beta_0\ep^2\pi''(1) q\p_Z q \\ 
        &+ \cancel{\beta_0\ep^2\pi'(1) \beta_\para\p_Zq} - \cancel{2}\beta_0\ep^2\pi'(1) q\p_Z q + \cancel{\beta_0\ep^2\pi'(1) q\p_Z\beta_\para} + O\pth{\ep^3}. 
\end{align*}
There are no explicit $O(1)$ contributions. Considering cancellations between different $S$ terms, the only $O(\ep)$ contributions are
\begin{align*}
    \ep S_0 &: - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi} \\ 
    \ep S_2 &: - \ep\br{\beta_\para, \Psi}. 
\end{align*}
At $O\pth{\ep^2}$, we have 
\begin{align*}
    \ep S_0 &: -\ep^2\nu_\para \p_Z\nu_\para \\ 
    \ep S_1 &: - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 = -\ep^2\np\psi \cdot \p_Z\np\psi = -\ep^2 \Braket{\psi, \p_Z\psi} \\ 
    \ep S_2 &: - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi}
\end{align*}

@@

Should I do $p=p_0\pi(r)$? This conflicts with $\ep=2\pi a/L$. Should use $\bm{e}_Z$ instead of $\bm{e}_z$ once I've nondimensionalized. Should also replace $\np^2$ with $\Delta$.

@@

Can I just split the pressures like $P'(\fr) = P'(1) + \cdots$

@@

Oh shit why is $|\nu_\perp|^2 = |\np\phi|^2 + |\np\psi|^2$? I should get $+2\br{\psi,\phi}$

I need to just write out the limit systems in their entirety. Probably need to just include pressure expansions.

>>>>

\title{Reduced MHD Notes}
\author{Finny Valorz}
%\date{June 2024}
\begin{document}
\maketitle

@@

More explicitly, we observe that $D_yf_0(x,y)$ is not invertible, failing to satisfy the constraint \eqref{constraint}:
\begin{equation} \begin{split}
    D_yf_0(x,y) [\delta y] = \pth{\deriv{f_0^j}{y^i} \br{\delta y^i}} 
    &= \pmat{\deriv{}{r}f_0^r & \deriv{}{\bm{\nu}_\perp}f_0^r & \deriv{}{\beta_\para}f_0^r\\ \deriv{}{r}f_0^{\bm{\nu}_\perp} & \ddots & \vdots \\ \deriv{}{r}f_0^{\beta_\para} & \cdots & \deriv{}{\beta_\para}f_0^{\beta_\para}} 
    \bmat{\delta r \\ \delta \bm{\nu}_\perp \\ \delta \beta_\para} \\
    &= \pmat{0 & -\np \cdot & 0 \\ 0 & 0 & -\np \\ 0 & -\np \cdot & 0} 
    \bmat{\delta r \\ \delta \bm{\nu}_\perp \\ \delta \beta_\para}
    = \pmat{-\np \cdot \delta \bm{\nu}_\perp \\ -\np \delta \beta_\para \\ -\np \cdot \delta \bm{\nu}_\perp }.
\end{split} \end{equation}
Example: $\deriv{f_0}{\bm{\nu}_\perp}[\delta\bm{\nu}_\perp] = \delta f_0 = \deriv{f_0}{\bm{\nu}_\perp}\cdot \delta\bm{\nu}_\perp = f_0(\bm{\nu}_\perp+ \delta\bm{\nu}_\perp) - f_0(\bm{\nu}_\perp) = \np\cdot \delta\bm{\nu}_\perp$. Also, $\frac{d}{d\ep}\np\cdot\pth{\bm{\nu}_\perp+\ep\delta\bm{\nu}_\perp} = \np\cdot\delta\bm{\nu}_\perp$.

Remember that partial derivatives are treated as in \eqref{partial} as $\p f_0^j/\p y^i \br{\delta y^i} = D_{y^i} f_0^j \br{\delta y^i}$. Regardless of the non-zero entries, this block-matrix has a column of zeroes, so its determinant is $0$ and it is non-invertible. {\color{red}It is probably best to avoid the notion of determinant of an operator between infinite-dimensional spaces. Invertibility is a dimension-independent notion, but determinant is not. --JB} Intuitively, a fast-slow system ought to change on the fast timescale with a change in one of its fast variables. The system above is degenerate in the fast variable $r$, which suggests there are hidden slow variables among these fast variables.

@@

!!!Here is how I should have done it!!!
\begin{align*}
    \p_\tau\nu_\para &= -\ep\nu_\para\p_Z\nu_\para - \bm{\nu}_\perp\cdot \np\nu_\para \\ 
        &- \frac{1}{M_0^2\beta_0} \fr \pth{\beta_0\ep \p_ZP\pth{\frinv} + \ep^2 \cancel{\pth{1+\ep\beta_\para}\p_Z\beta_\para} + \frac{1}{2}\ep^3\p_Z|\bm{\beta}_\perp|^2} \\ 
        &+ \frac{1}{M_0^2\beta_0} \fr\ep^2 \pth{\cancel{(1+\ep\beta_\para)\p_Z\beta_\para} + \bm{\beta}_\perp\cdot\np\beta_\para}. 
\end{align*}

@@

Equation \eqref{ndfaradayperp} gives an equation for the evolution of $\bm{\beta}_\perp = \np\Phi + \bm{e}_z\times\np\Psi$. 
\begin{align*} 
    \p_\tau\bm{\beta}_\perp = \p_Z \pth{\pth{1+\ep\beta_\para} \bm{\nu}_\perp - \ep\nu_\para \bm{\beta}_\perp} - \bm{e}_z \times \np \pth{\bm{e}_z \cdot \bm{\nu}_\perp \times \bm{\beta}_\perp}. 
\end{align*}
We use similar techniques to find $\Phi$ and $\Psi$ as we did for $\phi$ and $\psi$. The divergence of \eqref{ndfaradayperp} gives an expression for $\Phi$. 
\begin{align} \label{Phievolution} 
    \np^2\p_\tau\Phi = \np\cdot\p_\tau\bm{\beta}_\perp &= \p_Z \pth{\ep\np\beta_\para\cdot\bm{\nu}_\perp + (1+\ep\beta_\para) \np\cdot\bm{\nu}_\perp - \ep\np\nu_\para\cdot\bm{\beta_\perp} - \ep\nu_\para\np\cdot\bm{\beta_\perp}} \nonumber\\ 
    &= \p_Z\pth{\ep\Braket{\beta_\para, \phi} - \ep\br{\beta_\para, \psi} + (1+\ep\beta_\para) \np^2\phi - \ep\Braket{\nu_\para, \Phi} + \ep\br{\nu_\para, \Psi} -\ep\nu_\para\np^2\Phi} \nonumber\\ 
        &= \ep\Braket{\p_Z\beta_\para, \phi} + \ep\Braket{\beta_\para, \p_Z\phi} - \ep\br{\p_Z\beta_\para, \psi} - \ep\br{\beta_\para, \p_Z\psi} \nonumber\\ 
        &+ \ep\p_Z\beta_\para\np^2\phi + \pth{1+\ep\beta_\para}\np^2\p_Z\phi - \ep\Braket{\p_Z\nu_\para, \Phi} - \ep\Braket{\nu_\para, \p_Z\Phi} \nonumber\\ 
        &+ \ep\br{\p_Z\nu_\para, \Psi} + \ep\br{\nu_\para, \p_Z\Psi} - \ep\p_Z\nu_\para\np^2\Phi - \ep\nu_\para\np^2\p_Z\Phi. \text{\color{green}CHECKED} 
\end{align}
Note the second term in the equation of motion disappears because it involves the curl of a gradient.

@@

\subsection{$\nu_\para$ Evolution}
\begin{align} 
    \p_T\nu_\para &= \ep S_0 + \ep S_1 + \ep S_2 \text{, with}
\end{align}
\begin{align*}
    \ep S_0 &= -\ep^2\nu_\para \p_Z\nu_\para - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi}, \\ 
    \ep  S_1 &= -\beta_0\fr\p_Z\pi\pth{\frinv} - \cancel{\ep\p_Z\beta_\para} - \cancel{\ep^2q\p_Z\beta_\para} + \cancel{\ep^2\beta_\para\p_Z\beta_\para} \\ 
        &- \cancel{\ep^2\beta_\para\p_Z\beta_\para} - \frac{1}{2}\ep^2\p_Z |\np\Psi|^2 + O\pth{\ep^3}, \\ 
    \ep S_2 &= \cancel{\ep\p_Z\beta_\para} + \cancel{\ep^2 q\p_Z\beta_\para} - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} \\ 
        &- \ep\br{\beta_\para, \Psi} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi} + O\pth{\ep^3}. 
\end{align*}
Only one term changes in different regimes, namely 
\begin{align*}
    -\beta_0\fr\p_Z\pi\pth{\frinv} &= - \beta_0\ep\pi'(1) \p_Z\beta_\para + \beta_0\ep\pi'(1) \p_Zq \\ 
        &- \beta_0\ep^2\pi'(1) \pth{\cancel{q}-\beta_\para} \p_Z\beta_\para + \cancel{\beta_0\ep^2\pi'(1) \pth{q-\beta_\para} \p_Zq} \\ 
        &- \beta_0\ep^2\pi''(1) \beta_\para\p_Z \beta_\para + \beta_0\ep^2\pi''(1) \beta_\para\p_Zq + \beta_0\ep^2\pi''(1) q\p_Z\beta_\para - \beta_0\ep^2\pi''(1) q\p_Z q \\ 
        &+ \cancel{\beta_0\ep^2\pi'(1) \beta_\para\p_Zq} - \cancel{2}\beta_0\ep^2\pi'(1) q\p_Z q + \cancel{\beta_0\ep^2\pi'(1) q\p_Z\beta_\para} + O\pth{\ep^3}. 
\end{align*}
There are no explicit $O(1)$ contributions. Considering cancellations between different $S$ terms, the only $O(\ep)$ contributions are
\begin{align*}
    \ep S_0 &: - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi} \\ 
    \ep S_2 &: - \ep\br{\beta_\para, \Psi}. 
\end{align*}
At $O\pth{\ep^2}$, we have 
\begin{align*}
    \ep S_0 &: -\ep^2\nu_\para \p_Z\nu_\para \\ 
    \ep S_1 &: - \frac{1}{2}\ep^2\p_Z |\np\Psi|^2 = -\ep^2\np\Psi \cdot \p_Z\np\Psi = -\ep^2 \Braket{\Psi, \p_Z\Psi} \\ 
    \ep S_2 &: - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi}
\end{align*}

@@

@@

@@

I should make $t_0v_0/a=\ep$ from the get-go so that I don't have so many different time variables floating around.
